Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Environment

### 2015-06-01 Paper2frontiers/frontiers.tex
Updated with interns

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
Humans often interact with machines in uncertain and complicated environments, such as crowds and traffic, where they must contend with turbulence, moving obstacles, distractions, and disturbances. Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning. Sensory delays, sensory fusion, delays in sensory processing, and delays in action further confound our ability to behave reactively. The nervous system intelligently solves many problems by planning ahead \citep{belen1967control}, yet in the face of uncertainties we often cannot adequately prevent errors. There is the possibility, however, to exploit additional information from instruments that may be available -- particularly fast and accurate force sensors -- that can measure human machine interactions. Combining sensors with filtering techniques makes it possible to determine a person’s underlying intent -- the motion they would have done if they had not been disturbed. This intent provides new ways to understand the nature of control and provide novel feedback.

Recent work in our laboratory has attempted to outline the methodology for obtaining estimates of intent \citep{horowitz2015determining}. This method assumes a model of the dynamics and control of the human arm, and then through a manipulation of the equations of motion it integrates to find a unique estimate of intent. The algorithm recovers the trajectory a person intended to take, even if they were forced away from it due to environmental disturbances. This analysis has enabled us to show how subjects might sometimes alter their intent following exposure to unexpected force pulses.

A new question that arises is whether seeing one’s own intent, rather that what what actually happens, may be useful. The intent extraction method can be abriged to allow for real-time estimations and presented to the subject as a cursor. It is possible that performance using this estimated intention may compare quite favorably to the actual outcome in the presence of unexpected disturbances. If so, such a method holds great promise in any situation where humans and machines interact as it enables the machine to give the human operator what they want. This human-machine collaboration might be able to outperform what a person can do alone.

In contrast, displaying anything other than what truthfully happens is a distortion and a lie. Like many other visual distortion experiments \citep{miles1980long, pine1996learning}, intent feedback (IF) introduces a visuomotor discrepancy that may be confusing to the nervous system. Preplanning a specific route may not necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal. If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment without a specific intended route using optimal feedback control \citep{todorov2002optimal}. While this modeling strategy has been very successful at explaining data, it fundamentally assumes that corrective actions will be taken in response to relevant disturbances. Goals can also be reached at minimal \textit{expected} cost by constructing -- and possibly updating -- a specific intended route. IF may be troublesome if the nervous system cannot recognize it because no particular trajectory was intended.

While performance is the best indicator of IF's value, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want. Arm stiffness is known to increase during exposure to instability \citep{franklin2003adaptation} and uncertainty \citep{takahashi2001impedance}. We anticipated that these changes might also be modulated by presentation of IF. Reducing the effects of environmental instability and presenting the subject with a signal already known to them should make the arm relax and become more compliant. We hypothesized that any elevated arm stiffness from noisy disturbances would decrease when subjects received IF.

In this paper we describe our abridged method for real-time IF, and then expose subjects to IF in an unstable and unpredictable environment to compare their ability to perform goal-directed reaching using visual feedback of their hand position against visual feedback of their estimated intended position. We hypothesized that IF should lead to better performance in the presence of force-based disturbances. Accordingly, we hypothesized that during random disturbances, the intent trajectory should deviate less than the hand.

>>>>

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
Humans often interact with machines in uncertain and complicated environments, such as crowds and traffic, where they must contend with turbulence, moving obstacles, distractions, and disturbances. Despite our capacity to learn and adapt, some environments evolve too quickly or with too much uncertainty for meaningful learning. Sensory delays, sensory fusion, delays in sensory processing, and motor delays further confound our reactions. The nervous system intelligently solves many problems by planning ahead \citep{belen1967control}, yet in the face of uncertainties we often cannot adequately prevent errors. There is the possibility, however, to exploit additional information from instruments -- particularly fast and accurate force sensors -- that can measure human machine interactions. Combining sensors with filtering techniques makes it possible to determine a person’s underlying intent -- the motion they would have done had they not been disturbed. This intent provides new ways to understand the nature of control and provide novel feedback.

Recent work in our laboratory has attempted to outline a methodology for obtaining estimates of intent \citep{horowitz2015determining}. This method assumes a model of the dynamics and control of the human arm. Following manipulations of the equations of motion, the method integrates to find a unique estimate of intent. The algorithm recovers the trajectory a person intended to take, even if they were forced away from it due to environmental disturbances. This analysis has enabled us to show how some subjects sometimes alter their intent following exposure to unexpected force pulses.

A new question that arises is whether seeing one’s own intent, rather than what actually happens, may be useful. The intent extraction method can be streamlined to allow for real-time estimations of intent that can be presented to the subject as a cursor. Estimated intention may outperform the movement accuracy in the presence of unexpected disturbances. If so, such a method holds great promise in any situation where humans and machines interact as it enables the machine to give the human operator what they want. This human-machine collaboration could outperform what a person can do alone.

Displaying anything other than what truthfully happens is a distortion and a deceit. Like many other visual distortion experiments \citep{miles1980long, pine1996learning}, intent feedback (IF) introduces a visuomotor discrepancy that may be confusing to the nervous system. Preplanning a specific route may not be necessary, and instead the system might continuously react to the environmental disturbances until it reaches the goal. If people try to achieve a goal while minimizing some measure of cost, it is possible to compute a set of rules for reacting to the environment \citep{todorov2002optimal}. No specific intended route is needed when using optimal feedback control. While this modeling strategy has been very successful at explaining data, it fundamentally assumes that corrective actions will be taken in response to relevant disturbances. Goals can also be reached at minimal \textit{expected} cost by constructing -- and possibly updating -- a specific intended route. If no particular trajectory is intended, the nervous system could be unable to recognize IF.

While performance is the best indicator of IF's worth, changes in arm stiffness can provide supporting evidence that subjects are actually getting what they want. Arm stiffness is known to increase during exposure to instability \citep{franklin2003adaptation} and uncertainty \citep{takahashi2001impedance}. We anticipated that these changes might also be modulated by the presence of IF. Reducing the effects of environmental instability and presenting the subject with a signal already known to them should relax their arm and make it more compliant. We hypothesized that any elevated arm stiffness from noisy disturbances would decrease while subjects received IF.

In this paper we describe our streamlined method for real-time IF and then we expose subjects to IF in an unstable and unpredictable environment. We then compare their ability to perform goal-directed reaching using visual feedback of their hand position against IF. We hypothesized that IF should lead to better performance in the presence of force-based disturbances. Accordingly, we hypothesized that during random disturbances, the intent trajectory should deviate less than the hand trajectory.
